\begin{pa} \label{PA:4.5}
A student with a third floor dormitory window $32$ feet off the ground tosses a water balloon straight up in the air with an initial velocity of $16$ feet per second.  It turns out that the instantaneous velocity of the water balloon is given by the velocity function $v(t) = -32t + 16$, where $v$ is measured in feet per second and $t$ is measured in seconds.
\ba
\item Let $s(t)$ represent the height of the water balloon above the ground at time $t$, and note that $s$ is an antiderivative of $v$.  That is, $v$ is the derivative of $s$: $v(t) = s'(t)$.  Find a formula for $s(t)$ that satisfies the initial condition that the balloon is tossed from $32$ feet above ground.  In other words, make your formula for $s$ satisfy $s(0) = 32$.
\item At what time does the water balloon reach its maximum height?  At what time does the water balloon land?
\item Compute the three differences $s \left( \frac{1}{2} \right) - s(0)$, $s(2) - s \left( \frac{1}{2} \right)$, and $s(2) - s(0)$.  What do these differences represent?
\item What is the total vertical distance traveled by the water balloon from the time it is tossed until the time it lands?
\item Sketch a graph of the velocity function $y = v(t)$ on the time interval $[0,2]$.  What is the total net signed area bounded by $y = v(t)$ and the $t$-axis on $[0,2]$?  Answer this question in two ways:  first by using your work above, and then by using a familiar geometric formula to compute areas of certain relevant regions.
\ea
\end{pa} 

\afterpa