\begin{activity} \label{A:2.5.2}  
For each of the following functions, find the function's derivative.  State the rule(s) you use, label relevant derivatives appropriately, and be sure to clearly identify your overall answer. 
\bmtwo 
\ba
	\item $p(r) = 4\sqrt{r^6 + 2e^r}$
	\item $m(v) = \sin(v^2) \cos(v^3)$
	\item $\ds h(y) = \frac{\cos(10y)}{e^{4y}+1}$
	\item $s(z) = 2^{z^2 \sec (z)}$
	\item $\ds c(x) = \sin(e^{x^2})$
\ea
\emtwo
\end{activity}
\begin{smallhint}
\ba
	\item Use the constant multiple rule first, followed by the chain rule.
	\item Observe that $m$ is fundamentally a product of composite functions.
	\item Note that $h$ is a quotient of composite functions.
	\item The function $s$ is a composite function with outer function $2^z$.
	\item It is possible for a function to be a composite function with more than two functions in the chain.
\ea
\end{smallhint}
\begin{bighint}
\ba
	\item Note that $p'(r) = 4\frac{d}{dr}[\sqrt{r^6 + 2e^r}].$  Now use the chain rule to complete the remaining derivative.
	\item Observe that by the product rule, $m'(v) = \sin(v^2) \frac{d}{dv}[\cos(v^3)] + \cos(v^3) \frac{d}{dv}[\sin(v^2)].$  What rule is needed to execute the remaining two derivatives?
	\item Note that by the quotient rule,
	$$h'(y) = \frac{(e^{4y}+1) \frac{d}{dy}[\cos(10y)] - \cos(10y) \frac{d}{dy}[e^{4y}+1]}{(e^{4y}+1)^2}.$$
	What work remains to find $h'(y)$?
	\item By the chain rule, $s'(z) = 2^{z^2\sec(z)} \ln(2) \frac{d}{dz}[z^2 \sec(z)]$.  What rule should next be applied?
	\item If we first apply the chain rule to the outer function (the sine function), note that 
	$$c'(x) = \cos(e^{x^2}) \frac{d}{dx}[e^{x^2}].$$
	What rule is needed to differentiate the inner function, $e^{x^2}$?
\ea
\end{bighint}
\begin{activitySolution}
\ba
	\item By the constant multiple rule, $p'(r) = 4\frac{d}{dr}[\sqrt{r^6 + 2e^r}].$  Using the chain rule to complete the remaining derivative, we see that
	$$p'(r) = 4 \frac{1}{2\sqrt{r^6 + 2e^r}} \frac{d}{dr}[r^6 + 2e^r] = \frac{4(6r^5 + 2e^r)}{2\sqrt{r^6 + 2e^r}}.$$
	\item Observe that by the product rule, $m'(v) = \sin(v^2) \frac{d}{dv}[\cos(v^3)] + \cos(v^3) \frac{d}{dv}[\sin(v^2)].$  Applying the chain rule to differentiate $\cos(v^3)$ and $\sin(v^2)$, we see that 
	$$m'(v) = \sin(v^2) [-\sin(v^3) \cdot 3v^2] + \cos(v^3) [\cos(v^2) \cdot 2v] = -3v^2 \sin(v^2)\sin(v^3) + 2v \cos(v^3)\cos(v^2).$$ 
	\item By the quotient rule,
	$$h'(y) = \frac{(e^{4y}+1) \frac{d}{dy}[\cos(10y)] - \cos(10y) \frac{d}{dy}[e^{4y}+1]}{(e^{4y}+1)^2}.$$
	Applying the chain rule to differentiate $\cos(10y)$ and $e^{4y}$, it follows
	$$h'(y) = \frac{(e^{4y}+1) [-10\sin(10y)] - \cos(10y) [4e^{4y}]}{(e^{4y}+1)^2}.$$
	\item By the chain rule, $s'(z) = 2^{z^2\sec(z)} \ln(2) \frac{d}{dz}[z^2 \sec(z)]$.  Then by the product rule, we find that
	$$s'(z) = 2^{z^2\sec(z)} \ln(2) [z^2 \sec(z)\tan(z) + \sec(z) \cdot 2z].$$
	\item If we first apply the chain rule to the outer function (the sine function), note that 
	$$c'(x) = \cos(e^{x^2}) \frac{d}{dx}[e^{x^2}].$$
	Next, we again apply the chain rule, but this time to $e^{x^2}$, and get
	$$c'(x) = \cos(e^{x^2}) [e^{x^2}\cdot 2x].$$
\ea
\end{activitySolution}
\aftera