\begin{activity} \label{A:6.5.1}  Consider the following situations in which a varying force accomplishes work.
\ba
	\item Suppose that a heavy rope hangs over the side of a cliff.  The rope is 200 feet long and weighs 0.3 pounds per foot; initially the rope is fully extended.  How much work is required to haul in the entire length of the rope?  (Hint: set up a function $F(h)$ whose value is the weight of the rope remaining over the cliff after $h$ feet have been hauled in.)
	\item A leaky bucket is being hauled up from a 100 foot deep well.  When lifted from the water, the bucket and water together weigh 40 pounds.  As the bucket is being hauled upward at a constant rate, the bucket leaks water at a constant rate so that it is losing weight at a rate of 0.1 pounds per foot.  What function $B(h)$ tells the weight of the bucket after the bucket has been lifted $h$ feet?  What is the total amount of work accomplished in lifting the bucket to the top of the well?
	\item Now suppose that the bucket in (b) does not leak at a constant rate, but rather that its weight at a height $h$ feet above the water is given by $B(h) = 25 + 15e^{-0.05h}$.  What is the total work required to lift the bucket 100 feet?  What is the average force exerted on the bucket on the interval $h = 0$ to $h = 100$?
	\item From physics, \emph{Hooke's Law} \index{Hooke's Law} for springs states that the amount of force required to hold a spring that is compressed (or extended) to a particular length is proportionate to the distance the spring is compressed (or extended) from its natural length.  That is, the force to compress (or extend) a spring $x$ units from its natural length is $F(x) = kx$ for some constant $k$ (which is called the \emph{spring constant}.) For springs, we choose to measure the force in pounds and the distance the spring is compressed in feet.
	
	Suppose that a force of 5 pounds extends a particular spring 4 inches (1/3 foot) beyond its natural length.
	\be
		\item[i.]  Use the given fact that $F(1/3) = 5$ to find the spring constant $k$.
		\item[ii.]  Find the work done to extend the spring from its natural length to 1 foot beyond its natural length.
		\item[iii.]  Find the work required to extend the spring from 1 foot beyond its natural length to 1.5 feet beyond its natural length.
	\ee

\ea

\end{activity}
\begin{smallhint}
\ba
	\item Small hints for each of the prompts above.
\ea
\end{smallhint}
\begin{bighint}
\ba
	\item Big hints for each of the prompts above.
\ea
\end{bighint}
\begin{activitySolution}
\ba
	\item Solutions for each of the prompts above.
\ea
\end{activitySolution}
\aftera