\begin{activity} \label{A:7.6.1}  
  Our first model will be based on the following assumption:
  \begin{quote}{\em
      The rate of change of the population is proportional to the
      population.  
    }
  \end{quote}

  On the face of it, this seems pretty reasonable.  When there is a
  relatively small number of people, there will be fewer births and
  deaths so the rate of change will be small.  When there is a larger
  number of people, there will be more births and deaths so we expect
  a larger rate of change.

  If $P(t)$ is the population $t$ years after the year 2000, we may
  express this assumption as
  $$
  \frac{dP}{dt} = kP
  $$
  where $k$ is a constant of proportionality.

\ba
\item Use the data in the table to estimate the derivative $P'(0)$
  using a central difference.  Assume that $t=0$ corresponds to the
  year 2000.

\item What is the population $P(0)$?

\item Use these two facts to estimate the constant of proportionality
  $k$ in the differential equation.

\item Now that we know the value of $k$, we have the initial
  value problem
  $$
    \frac{dP}{dt} = kP, \ P(0) = 6.084.
  $$
  Find the solution to this initial value problem.

\item What does your solution predict for the population in the year
  2010?  Is this close to the actual population given in the table?

\item When does your solution predict that the population will reach
  12 billion?

\item What does your solution predict for the population in the year
  2500? 

\item Do you think this is a reasonable model for the earth's
  population?  Why or why not?  Explain your thinking using a couple
  of complete sentences. 

\ea
\end{activity}
\begin{smallhint}
\ba
	\item Small hints for each of the prompts above.
\ea
\end{smallhint}
\begin{bighint}
\ba
	\item Big hints for each of the prompts above.
\ea
\end{bighint}
\begin{activitySolution}
\ba
	\item Solutions for each of the prompts above.
\ea
\end{activitySolution}
\aftera