\begin{activity} \label{8.2.Act3} The formulas we have derived for the geometric series and its partial sum so far have assumed we begin indexing our sums at $n=0$. If instead we have a sum that does not begin at $n=0$, we can factor out common terms and use our established formulas.  This process is illustrated in the examples in this activity.
\ba
\item Consider the sum
\[ \sum_{k=1}^{\infty} (2)\left(\frac{1}{3}\right)^k = (2)\left(\frac{1}{3}\right) + (2)\left(\frac{1}{3}\right)^2 + (2)\left(\frac{1}{3}\right)^3 + \cdots .\]
Remove the common factor of $(2)\left(\frac{1}{3}\right)$ from each term and hence find the sum of the series. 

\item Next let $a$ and $r$ be real numbers with $-1<r<1$. Consider the sum  
\[\sum_{k=3}^{\infty} ar^k = ar^3+ar^4+ar^5 + \cdots.\]
Remove the common factor of $ar^3$ from each term and find the sum of the series. 

\item Finally, we consider the most general case. Let $a$ and $r$ be real numbers with $-1<r<1$, let $n$ be a positive integer, and consider the sum
\[ \sum_{k=n}^{\infty} ar^k = ar^n+ar^{n+1}+ar^{n+2} + \cdots .\]
Remove the common factor of $ar^n$ from each term to find the sum of the series. 

\ea
\end{activity}

\begin{smallhint}
\ba
	\item Small hints for each of the prompts above.
\ea
\end{smallhint}
\begin{bighint}
\ba
	\item Big hints for each of the prompts above.
\ea
\end{bighint}
\begin{activitySolution}
\ba
	\item Factoring out $(2)\left(\frac{1}{3}\right)$ gives us
\begin{align*}
(2)\left(\frac{1}{3}\right) \left[1 + \left(\frac{1}{3}\right) + \left(\frac{1}{3}\right)^2 + \cdots \right] &= (2)\left(\frac{1}{3}\right)\sum_{k=0}^{\infty} \left(\frac{1}{3}\right)^k \\
    &= (2)\left(\frac{1}{3}\right) \left(\frac{1}{1-\frac{1}{3}}\right) \\
    &= (2)\left(\frac{1}{3}\right) \left(\frac{3}{2}\right)\\
    &=  4.
\end{align*}
    \item Factoring out $ar^3$ gives us
\begin{align*}
ar^3+ar^4+ar^5 + \cdots &= ar^3\left(1+r+r^2+ \cdots \right) \\
    &= ar^3\sum_{k=0}^{\infty} r^k \\
    &= ar^3 \left(\frac{1}{1-r}\right) \\
    &= \frac{ar^3}{1-r}.
\end{align*}
    \item Factoring out $ar^n$ gives us
\begin{align*}
ar^n+ar^{n+1}+ar^{n+2} + \cdots &= ar^n\left(1+r+r^2+ \cdots \right) \\
    &= ar^n\sum_{k=0}^{\infty} r^k \\
    &= ar^n \left(\frac{1}{1-r}\right) \\
    &= \frac{ar^n}{1-r}.
\end{align*}
\ea
\end{activitySolution}
\aftera 