\begin{activity} \label{8.3.Act5} The series $\sum \frac{1}{k^p}$ are special series called $p$-series. We have already seen that the $p$-series with $p=1$ (the harmonic series) diverges. We investigate the behavior of other $p$-series in this activity.
\ba
    \item Evaluate the improper integral $\ds \int_1^{\infty} \frac{1}{x^2} \ dx$. Does the series $\ds \sum_{k=1}^{\infty} \frac{1}{k^2}$
converge or diverge? Explain.


    \item Evaluate the improper integral $\ds \int_1^{\infty} \frac{1}{x^p} \ dx$ where $p > 1$. For which values of $p$ can we conclude that the series $\ds \sum_{k=1}^{\infty} \frac{1}{k^p}$
converges? 

    \item Evaluate the improper integral $\ds \int_1^{\infty} \frac{1}{x^p} \ dx$ where $p < 1$. What does this tell us about the corresponding $p$-series
$\ds \sum_{k=1}^{\infty} \frac{1}{k^p}$?

\item Summarize your work in this activity by completing the following statement.
\begin{quote}
The $p$-series $\ds \sum_{k=1}^{\infty} \frac{1}{k^p}$ converges if and only if \underline{\hspace{2in}}.
\end{quote}

\ea
\end{activity}

\begin{smallhint}
\ba
	\item Small hints for each of the prompts above.
\ea
\end{smallhint}
\begin{bighint}
\ba
	\item Big hints for each of the prompts above.
\ea
\end{bighint}
\begin{activitySolution}
\ba
	\item We evaluate the improper integral:
\begin{align*}
\int_{1}^{\infty} \frac{1}{x^2} \ dx &= \lim_{t \to \infty} \int_{1}^{t} \frac{1}{x^2} \ dx \\
    &= \lim_{t \to \infty} -\frac{1}{x} \biggm|_{1}^{t} \\
    &= \lim_{t \to \infty} \left(-\frac{1}{t} + \frac{1}{1} \right) \\
    &= 1.
\end{align*}
So the improper integral converges. The Integral Test then shows that the series $\ds \sum_{k=1}^{\infty} \frac{1}{k^2}$ converges.
    \item Assume $p > 1$. Then $p-1 > 0$ and so $x^{-p+1} = \frac{1}{x^{p-1}}$ and
\[\ds \lim_{x \to \infty} x^{-p+1} = \lim_{x \to \infty} \frac{1}{x^{p-1}} = 0.\]
Thus,
\begin{align*}
\int_{1}^{\infty} \frac{1}{x^p} \ dx &= \lim_{t \to \infty} \int_{1}^{t} \frac{1}{x^p} \ dx \\
    &= \lim_{t \to \infty} \frac{x^{-p+1}}{-p+1} \biggm|_{1}^{t} \\
    &= \frac{1}{1-p} \lim_{t \to \infty} \left(t^{-p+1} -  1 \right) \\
    &= \frac{1}{p-1}.
\end{align*}
So the improper integral $\ds \int_1^{\infty} \frac{1}{x^p} \ dx$ converges when $p > 1$. The Integral Test then shows that the series $\ds \sum_{k=1}^{\infty} \frac{1}{k^p}$ converges when $p > 1$.
    \item Assume $p < 1$. Then $1-p > 0$ and so
\[\ds \lim_{x \to \infty} x^{-p+1} = \infty.\]
Thus,
\begin{align*}
\int_{1}^{\infty} \frac{1}{x^p} \ dx &= \lim_{t \to \infty} \int_{1}^{t} \frac{1}{x^p} \ dx \\
    &= \lim_{t \to \infty} \frac{x^{-p+1}}{-p+1} \biggm|_{1}^{t} \\
    &= \frac{1}{1-p} \lim_{t \to \infty} \left(t^{-p+1} -  1 \right) \\
    &= \infty.
\end{align*}
So the improper integral $\ds \int_1^{\infty} \frac{1}{x^p} \ dx$ diverges when $p < 1$. The Integral Test then shows that the series $\ds \sum_{k=1}^{\infty} \frac{1}{k^p}$ also diverges when $p < 1$.
    \item We complete the statement as
\begin{quote}
The $p$-series $\ds \sum_{k=1}^{\infty} \frac{1}{k^p}$ converges if and only if $p > 1$.
\end{quote}
\ea
\end{activitySolution}
\aftera 