\begin{adjustwidth*}{}{-2.25in}
\textbf{{\large Exercises}}
\setlength{\columnsep}{25pt}
\begin{multicols*}{2}
\noindent Terms and Concepts \small
\begin{enumerate}[1)]
\item T/F: Given a function $f(x)$, Newton's Method produces an exact solution to $f(x) = 0$.
\item T/F: In order to get a solution to $f(x)=0$ accurate to $d$ places after the decimal, at least $d+1$ iterations of Newton's Method must be used.
\end{enumerate} 

\noindent {\normalsize Problems} \small

\noindent{\bf In exercises 3--7, the roots of $f(x)$ are known or are easily found. Use 5 iterations of Newton's Method with the given initial approximation to approximate the root. Compare it to the known value of the root.}

\begin{enumerate}[1),resume]
\item $f(x) = \cos x$, $x_0=1.5$
\item $f(x) = \sin x$, $x_0=1$
\item $f(x) = x^2+x-2$, $x_0=0$
\item $f(x) = x^2-2$, $x_0=1.5$
\item $f(x) = \ln x$, $x_0=2$
\end{enumerate}

\noindent{\bf In exercises 8--11, use Newton's Method to approximate all roots of the given functions accurate to 3 places after the decimal.
If an interval is given, find only the roots that lie in that interval.}

\begin{enumerate}[1),resume]
\item $f(x) = x^3+5x^2-x-1$
\item $f(x) = x^4+2x^3-7x^2-x+5$
\item $f(x) = x^{17}-2x^{13}-10x^8+10$ on $(-2,2)$
\item $f(x) = x^2\cos x + (x-1)\sin x$ on $(-3,3)$
\end{enumerate}

\noindent{\bf In exercises 12--15, use Newton's Method to approximate when the given functions are equal, accurate to 3 places after the decimal.}

\begin{enumerate}[1),resume]
\item $f(x) = x^2$, $g(x) = \cos x$
\item $f(x) = x^2-1$, $g(x) = \sin x$
\item $f(x) = e^{x^2}$, $g(x) = \cos x$
\item $f(x) = x$, $g(x) = \tan x$ on $[-6,6]$

\item Why does Newton's Method fail in finding a root of $f(x) = x^3-3x^2+x+3$ when $x_0=1$?

\item Why does Newton's Method fail in finding a root of $f(x) = -17x^4+130x^3-301x^2+156x+156$ when $x_0=1$?

\item A rectangular piece of cardboard of dimensions $8\times 17$
is used to make an open-top box by cutting out a small square of side
$x$ from each corner and bending up the sides. 
If $x=2$, then the volume of the box is $2\cdot 4\cdot 13=104$. Use
Newton's method to find a value of $x$ for which the box has volume
100, accurate to 3 significant figures.
%\begin{answer} $2.19$ or $1.26$
%\end{answer}

\end{enumerate}

%------------------------------------------
% END OF EXERCISES ON FIRST PAGE
%------------------------------------------
\end{multicols*}
\end{adjustwidth*}

\afterexercises