\begin{example}
For a falling ball whose position function is given by \newline $s(t) = 16 - 16t^2$ where $s$ is measured in feet and $t$ in seconds, find an expression for the average velocity of the ball on a time interval of the form $[0.5, 0.5+h]$ where $-0.5 < h < 0.5$ and $h \ne 0$.  Use this expression to compute the average velocity on $[0.5,0.75]$ and $[0.4,0.5]$, as well as to make a conjecture about the instantaneous velocity at $t = 0.5$.

\solution We make the assumptions that $-0.5 < h < 0.5$ and $h \ne 0$ because $h$ cannot be zero (otherwise there is no interval on which to compute average velocity) and because the function only makes sense on the time interval $0 \le t \le 1$, as this is the duration of time during which the ball is falling.  Observe that we want to compute and simplify $$AV_{[0.5, 0.5+h]} = \frac{s(0.5+h) - s(0.5)}{(0.5+h) - 0.5}.$$  The most unusual part of this computation is finding $s(0.5+h)$.  To do so, we follow the rule that defines the function $s$.  In particular, since $s(t) = 16-16t^2$, we see that 
\begin{eqnarray*}
s(0.5+h) & = & 16 - 16(0.5 + h)^2 \\
& = & 16 - 16(0.25 + h + h^2) \\
& = & 16 - 4 - 16h - 16h^2 \\
& = & 12 - 16h - 16h^2.
\end{eqnarray*}
Now, returning to our computation of the average velocity, we find that 
\begin{eqnarray*}
 AV_{[0.5, 0.5+h]} & = & \frac{s(0.5+h) - s(0.5)}{(0.5+h) - 0.5} \\
& = & \frac{(12 - 16h - 16h^2) - (16 - 16(0.5)^2)}{0.5 + h - 0.5} \\
& = & \frac{12 - 16h - 16h^2 - 12}{h} \\
& = & \frac{-16h - 16h^2}{h}.
\end{eqnarray*}
At this point, we note two things:  first, the expression for average velocity clearly depends on $h$, which it must, since as $h$ changes the average velocity will change.  Further, we note that since $h$ can never equal zero, we may further simplify the most recent expression.  Removing the common factor of $h$ from the numerator and denominator, it follows that
$$ AV_{[0.5, 0.5+h]} = -16 - 16h.$$
Now, for any small positive or negative value of $h$, we can compute the average velocity.  For instance, to obtain the average velocity on $[0.5,0.75]$, we let $h = 0.25$, and the average velocity is $-16 - 16(0.25) = -20$ ft/sec.  To get the average velocity on $[0.4, 0.5]$, we let $h = -0.1$, which tells us the average velocity is $-16 - 16(-0.1) = -14.4$ ft/sec.  Moreover, we can even explore what happens to $AV_{[0.5, 0.5+h]}$ as $h$ gets closer and closer to zero.  As $h$ approaches zero, $-16h$ will also approach zero, and thus it appears that the instantaneous velocity of the ball at $t = 0.5$ should be $-16$ ft/sec.
\end{example}