\begin{example} \label{Ex:2.1.Eg4}
A company manufactures rope, and the total cost of producing $r$ feet of rope is $C(r)$ dollars.  What does it mean to say that $C(2000) = 800$? What are the units of $C'(r)$? Suppose that $C(2000) = 800$ and $C'(2000) = 0.35$.  Estimate $C(2100)$, and justify your estimate by writing at least one sentence that explains your thinking. Which of the following statements do you think is true, and why?
	\begin{itemize}
		\item $C'(2000) < C'(3000)$
		\item $C'(2000) = C'(3000)$
		\item $C'(2000) > C'(3000)$
	\end{itemize}
Suppose someone claims that $C'(5000) = -0.1$.  What would the practical meaning of this derivative value tell you about the approximate cost of the next foot of rope?  Is this possible?  Why or why not?  
	
\solution When we say $C(2000) = 800$, we mean that the total cost of producing $2000$ feet of rope is $800$ dollars.  Remember that the units on any derivative are ``units of output per unit of input,'' and the units of $C'(r)$ are ``dollars per foot.''

If $C(2000) = 800$ and $C'(2000) = 0.35$, then we know once $2000$ feet of rope are produced, the total cost function is increasing at $\$0.35$ per additional foot of rope.  Then, if we manufacture an additional $100$ feet of rope, the additional total cost will be approximately 
\[ 100 \ \mbox{feet} \cdot 0.35 \ \frac{\mbox{dollars}}{\mbox{foot}} = 35 \ \mbox{dollars}. \]
Therefore, we find that $C(2100) \approx C(2000) + 35 = 835,$ or that the cost to make $2100$ feet of rope is about $\$835$.

Either $C'(2000) = C'(3000)$ or $C'(2000) > C'(3000)$, since we expect the cost per foot of additional rope to either stay constant or to get smaller as the production volume increases.  Said differently, the instantaneous rate of change of the total cost function should either be constant or decrease due to economy of scale.

It is impossible to have $C'(5000) = -0.1$ and indeed to have any negative derivative value for the total cost function.  The total cost function $C(r)$ can never decrease, because it doesn't make sense for the total cost of producing $5001$ feet of rope to be less than the total cost of producing $5000$ feet of rope.

\end{example}