%\begin{marginfigure}[4cm] % MARGIN FIGURE
%\subfloat[]{\margingraphics{figures/figseries2a}}
%\subfloat[]{\margingraphics{figures/figseries2b}}
%\subfloat[]{\margingraphics{figures/figseries2c}}
%\caption{Scatter plots relating to Example \ref{eg:7.2.2}.} \label{F:eg:7.2.2}
%\end{marginfigure}

\begin{example} \label{eg:7.2.3} % EXAMPLE
Determine the convergence of the following series.\\

\noindent\begin{minipage}[t]{.33\linewidth}
\begin{enumerate}
\item		$\ds\sum_{n=1}^\infty \frac{1}{n}$
\item		$\ds\sum_{n=1}^\infty \frac{1}{n^2}$
\end{enumerate}
\end{minipage}
\begin{minipage}[t]{.33\linewidth}
\begin{enumerate}\addtocounter{enumi}{2}
\item		$\ds\sum_{n=1}^\infty \frac{1}{\sqrt{n}}$
\item		$\ds\sum_{n=1}^\infty \frac{(-1)^n}{n}$
\end{enumerate}
\end{minipage}\begin{minipage}[t]{.33\linewidth}
\begin{enumerate}\addtocounter{enumi}{4}
\item		$\ds\sum_{n=10}^\infty \frac{1}{(\frac12n-5)^3}$
\item		$\ds\sum_{n=1}^\infty \frac{1}{2^n}$
\end{enumerate}
\end{minipage} 

\solution
\begin{enumerate}
\item		This is a $p$-series with $p=1$. By Theorem \ref{thm:pseries}, this series diverges. 

This series is a famous series, called the \emph{Harmonic Series}, so named because of its relationship to \emph{harmonics} in the study of music and sound. 

\item		This is a $p$-series with $p=2$. By Theorem \ref{thm:pseries}, it converges. Note that the theorem does not give a formula by which we can determine \emph{what} the series converges to; we just know it converges. A famous, unexpected result is that this series converges to $\ds\frac{\pi^2}{6}$.

\item		This is a $p$-series with $p=1/2$; the theorem states that it diverges.

\item		This is not a $p$-series; the definition does not allow for alternating signs. Therefore we cannot apply Theorem \ref{thm:pseries}. (Another famous result states that this series, the \emph{Alternating Harmonic Series}, converges to $\ln 2$.)

\item		This is a general $p$-series with $p=3$, therefore it converges.

\item		This is not a $p$-series, but a geometric series with $r=2$. It converges.
\end{enumerate}
\end{example}