\section{Trigonometric Integrals} \label{S:5.2.TrigInt}

\begin{goals}
\item How do we integrate functions that are products of powers of the trigonometric functions?
\item How do we integrate functions that are products of trigonometric functions whose angles are constant multiples of a single variable?
\end{goals}

%-----------------------------------
%SUBSECTION: INTRODUCTION
%-----------------------------------

\subsection*{Introduction}

Functions involving trigonometric functions are useful for describing periodic behavior.  Learning techniques to integrate such functions helps solve problems involving periodic behavior and build general problem solving skills.
 This section describes several techniques for finding antiderivatives of certain combinations of trigonometric functions.

In Preview Activity~\ref{PA:5.2} we refresh our use of substitution and integration by parts to investigate some indefinite integrals that involve products of trigonometric functions.

\input{previews/5.2.PA1} %PREVIEW 

%-----------------------------------------------------------------
% SUBSECTION INTEGRALS OF THE FORM $\ds \int \sin^m x\cos^n x\ dx$
%-----------------------------------------------------------------
\subsection*{Integrals of the form $\ds \int \sin^m (x) \cos^n (x)\ dx$} 

In Preview Activity~\ref{PA:5.2}, we integrate $\int \sin(x) \cos(x) \ dx$; the integration is not difficult, and one could easily evaluate the indefinite integral by letting $u=\sin(x) $ or by letting $u = \cos(x)$. This integral is easy since the power of both sine and cosine is $1$. 

What if the power of both sine and cosine are not $1$? Preview Activity~\ref{PA:5.2} gives an idea on how to use substitution to evaluate such integrals. We generalize the integral $\int \sin(x) \cos(x) \ dx$ and consider integrals of the form $\int \sin^m(x) \cos^n(x)\ dx$, where $m$ and $n$ are nonnegative integers. Many times we use the Pythagorean identity $\cos^2 (x) + \sin^2(x) = 1$ to convert high powers of one trigonometric function into the other while leaving a single sine or cosine term in the integrand. This technique will allow us to use substitution to evaluate such integrals, and we summarize it in the following.

\concept{Integrals Involving Powers of Sine and Cosine} %CONCEPT
{Consider $\ds \int \sin^m(x) \cos^n(x) \ dx$, where $m$ and $n$ are nonnegative integers.\index{integration!of trig. powers}
\begin{enumerate}[1)]
\item If $m$ is odd, then $m=2k+1$ for some integer $k$. Rewrite \\
$$ \sin^m(x) = \sin^{2k+1}(x) = \sin^{2k}(x)\sin(x) $$
$$= (\sin^2(x))^k\sin(x) = (1-\cos^2(x))^k\sin(x).$$ 
Then 
$$\int \sin^m(x)\cos^n(x)\ dx $$
$$= \int (1-\cos^2(x))^k\sin(x)\cos^n(x)\ dx = -\int (1-u^2)^ku^n\ du,$$ 
where $u = \cos(x)$ and $du = -\sin(x)\ dx$. 

\item If $n$ is odd, then using substitutions similar to that outlined above we have
$$ \int \sin^m(x)\cos^n(x)\ dx = \int u^m(1-u^2)^k\ du,$$ %
where $u = \sin(x)$ and $du = \cos(x)\ dx$.

\item If both $m$ and $n$ are even, use the power--reducing identities %\small
$$\cos^2(x) = \frac{1+\cos (2x)}{2} \quad \text{and}\quad \sin^2(x) = \frac{1-\cos(2x)}2$$ %\normalsize
to reduce the degree of the integrand. Expand the result and apply the principles again.
\end{enumerate}
} %END CONCEPT

To demonstrate, we consider the following examples.

\input{examples/5-2_Eg1} % EXAMPLE

\marginnote[-4cm]{\textbf{Technology Note:} The work we are doing here can be a bit tedious, but the skills it develops (problem solving, algebraic manipulation, etc.) are important. Nowadays problems of this sort are often solved using a computer algebra system. The powerful program \textit{WolframAlpha}\textsuperscript{\textregistered} integrates $\int \sin^5(x)\cos^9(x)\ dx$ as 
\begin{align*}
f(x)&=-\frac{45 \cos (2 x)}{16384}-\frac{5 \cos (4 x)}{8192}+\frac{19 \cos (6x)}{49152}\\
&+\frac{\cos (8 x)}{4096}-\frac{\cos (10 x)}{81920}-\frac{\cos (12x)}{24576}-\frac{\cos (14 x)}{114688},
\end{align*}
which clearly has a different form than our answer in Example \ref{eg:5.2.1}, which is
\begin{align*}
g(x)&=\frac16\sin^6 (x)-\frac12\sin^8 (x)+\frac35\sin^{10} (x)\\
&-\frac13\sin^{12} (x)+\frac{1}{14}\sin^{14} (x).
\end{align*}
Figure \ref{F:5.2_Eg2} shows a graph of $f$ and $g$; they are clearly not equal. We leave it to the reader to recognize why both answers are correct.}

\input{examples/5-2_Eg2} % EXAMPLE

\input{examples/5-2_Eg3} % EXAMPLE

The process above was a bit long and tedious, but being able to work a problem such as this from start to finish is important.

\input{activities/5.2.Act1} %ACTIVITY  

%------------------------------------------------
% SUBSECTION Integrals of the form $\ds \int\sin(mx)\sin(nx)\ dx,$ $\ds\int \cos(mx)\cos(nx)\ dx$, and $\ds\int \sin(mx)\cos(nx)\ dx$
%------------------------------------------------
\subsection*{Integrals of the form $\ds \int\sin(mx)\sin(nx)\ dx,$ $\ds\int \cos(mx)\cos(nx)\ dx$, and $\ds\int \sin(mx)\cos(nx)\ dx$}

Functions that contain products of sines and cosines of differing periods are important in many applications including the analysis of sound waves. Integrals of the form \small
$$\int\sin(mx)\sin(nx)\ dx,\quad \int \cos(mx)\cos(nx)\ dx \quad \text{and}\quad\int \sin(mx)\cos(nx)\ dx$$ \normalsize
are best approached by first applying the Product to Sum Formulas, namely
\begin{align*}
\sin(mx)\sin(nx) &= \frac12\Big[\cos\big((m-n)x\big)-\cos\big((m+n)x\big)\Big] \\
\cos(mx)\cos(nx) &= \frac12\Big[\cos\big((m-n)x\big)+\cos\big((m+n)x\big)\Big] \\
\sin(mx)\cos(nx) &=	\frac12\Big[\sin\big((m-n)x\big)+\sin\big((m+n)x\big)\Big]
\end{align*}

\input{examples/5-2_Eg6} % EXAMPLE

%------------------------------------------------------
% SUBSECTION Integrals of the form $\ds\int\tan^mx\sec^nx\ dx$
%------------------------------------------------------
\subsection*{Integrals of the form $\ds\int\tan^m(x)\sec^n(x)\ dx$} 

When evaluating integrals of the form $\int \sin^m(x)\cos^n(x)\ dx$, the Pythagorean Theorem allowed us to convert even powers of sine into even powers of cosine, and vice--versa. If, for instance, the power of sine was odd, we pulled out one $\sin(x)$ and converted the remaining even power of $\sin(x)$ into a function using powers of $\cos(x)$, leading to an easy substitution.

The same basic strategy applies to integrals of the form \newline $\int \tan^m(x)\sec^n(x)\ dx$, albeit a bit more nuanced. The following three facts will prove useful:
\begin{itemize}
\item $\frac{d}{dx}(\tan(x)) = \sec^2(x)$, 
\item $\frac{d}{dx}(\sec(x)) = \sec(x)\tan(x)$ , and 
\item	$1+\tan^2(x) = \sec^2(x)$ (the Pythagorean Theorem).
\end{itemize}

If the integrand can be manipulated to separate a $\sec^2(x)$ term with the remaining secant power even, or if a $\sec(x)\tan(x)$ term can be separated with the remaining $\tan(x)$ power even, the Pythagorean Theorem can be employed, leading to a simple substitution. This strategy is outlined in the following:

\concept{Integrals Involving Powers of Tangent and Secant} % CONCEPT
{Consider $\ds\int\tan^m(x)\sec^n(x)\ dx$, where $m$ and $n$ are nonnegative integers.\index{integration!of trig. powers}
\begin{enumerate}[1)]
\item If $n$ is even, then $n=2k$ for some integer $k$. Rewrite $\sec^n(x)$ as 
$$\sec^n(x) = \sec^{2k}(x) = \sec^{2k-2}(x)\sec^2(x)$$
$$ = (1+\tan^2(x))^{k-1}\sec^2(x).$$
Then
$$\int\tan^m(x)\sec^n(x)\ dx=$$
$$\int\tan^m(x)(1+\tan^2(x))^{k-1}\sec^2(x)\ dx =$$
$$ \int u^m(1+u^2)^{k-1}\ du,$$
where $u = \tan(x)$ and $du = \sec^2(x)\ dx$.

\item If $m$ is odd, then $m=2k+1$ for some integer $k$. Rewrite $\tan^m(x)\sec^n(x)$ as
$$\tan^m(x)\sec^n(x) = \tan^{2k+1}(x)\sec^n(x) = $$
$$\tan^{2k}(x)\sec^{n-1}(x)\sec(x)\tan(x) = $$
$$(\sec^2(x)-1)^k\sec^{n-1}(x)\sec(x)\tan(x).$$
Then
$$\int\tan^m(x)\sec^n(x)\ dx=$$
$$\int(\sec^2(x)-1)^k\sec^{n-1}(x)\sec(x)\tan(x)\ dx = $$
$$\int(u^2-1)^ku^{n-1}\ du,$$
where $u = \sec(x)$ and $du = \sec(x)\tan(x)\ dx$.

\item If $n$ is odd and $m$ is even, then $m=2k$ for some integer $k$. Convert $\tan^m(x) $ to $(\sec^2(x)-1)^k$. Expand the new integrand and use Integration By Parts, with $dv = \sec^2(x)\ dx$.

\item If $m$ is even and $n=0$, rewrite $\tan^m(x)$ as
$$\tan^m(x) = \tan^{m-2}(x)\tan^2(x) = \tan^{m-2}(x)(\sec^2(x)-1) = $$
$$\tan^{m-2}(x)\sec^2(x)-\tan^{m-2}(x).$$
So
$$\int\tan^m(x)\ dx = $$
$$\underbrace{\int\tan^{m-2}(x)\sec^2(x)\ dx}_{\text{\small apply rule \#1}}\quad - \underbrace{\int\tan^{m-2}(x)\ dx}_{\text{\small apply rule \#4 again}}.$$

\end{enumerate}
} %end concept

The techniques described in items $1$) and $2$) above are relatively straightforward, but the techniques in items $3$) and $4$) can be rather tedious. A few examples will help with these methods.

\input{examples/5-2_Eg4} % EXAMPLE

\input{examples/5-2_Eg5} % EXAMPLE  

\input{activities/5.2.Act2} % ACTIVITY 

%------------------------------------------------------
% SUBSECTION Integrals of the form $\ds\int\csc^mx\cot^nx\ dx$
%------------------------------------------------------
\subsection*{Integrals of the form $\ds\int\csc^m(x)\cot^n(x)\ dx$} 

We can use similar methods to antidifferentiate integrals involving powers of cosecant and cotangent.  We leave the development of the process and practice to the reader.

%----------------------------------------------------------
% SUMMARY
%----------------------------------------------------------
\begin{summary}
\item The method of Trigonometric Integration enables functions of the form $\ds \sin^m (x) \cos^n (x)$, $\ds \sin(mx)\sin(nx)$, $\ds \cos(mx) \cos(nx)$, and $\ds (mx)\cos(nx)$ to be antidifferentiated.
\item Also using this method, we can antidifferentiate functions of the form $\ds \tan^m(x)\sec^n(x)$ and $\csc^m(x)\cot^n(x)$.
\end{summary}

\clearpage

%--------------
% EXERCISES
%--------------
\input{exercises/5-2.TrigInt(Ex)} 

\cleardoublepage